Inferring neural networks from fluorescent calcium imaging using partial correlation and sample weighting

### 2014-06-05 article.tex
Motivation for directivity

Our first guess was that introduce an ensemble method would improve performance
[cite papers which praise ensemble methods]. Among all parameters of our method,
we choose to average over two parameters, i.e. over multiple combinations of the
hard-threshold value $f_3$ and the pattern of the low-pass filter $c_l$. In order to 
free ourself from the choice of a probability distribution for the $(f_3,c_l)$ combinations,
we choose a uniform distribution for the threshold distribution (in a given interval of 
values) within each filtering process and perform a weighted average for the low-pass 
filter pattern. Formally, the stacked predicted adjacency matrix can be formulated by

@@

\section{Causal discovery methodology}
% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since the partial correlation is a symmetric, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both ) can't
be inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.} \cite{de2004discovery}.
In our last submission in hope to improve performance, some causal information
(directivity of the links) were added by stacking the partial correlation score
matrix with an pairwise activation heuristics. This method computes an anti-symmetric
activation score $z_{ij}$ based the variation of fluorescence signal of a neuron
$j$ due to a neuron $i$ over all timesteps:
\[
z_{ij} = \sum_{t=1}^n I(X_{t+1,j} - X_{t, i} \in [f_1, f_2]) -  I(X_{t+1,i} - X_{t, j} \in [f_1, f_2])
\]
where $f_1$ and $f_2$ are parameters of the method and $I$ is the
indicator function. % TODO say that we use only the 3 first filters for this

@@

\end{tabular}
\end{table}
\subsection{AUPRC}

>>>>

Our first guess was that introduce an ensemble method would improve performance
[cite papers which praise ensemble methods]. Among all parameters of our method,
we choose to average over two parameters, i.e. over multiple combinations of the
hard-threshold value $f_3$ and the pattern of the low-pass filter $c_l$. In order to 
free ourself from the choice of a probability distribution for the $(f_3,c_l)$ combinations,
we choose a uniform distribution for the threshold distribution (in a given interval of 
values) within each filtering process and perform a weighted average for the low-pass 
filter pattern. Formally, the stacked predicted adjacency matrix $y_{stacking}$ can be formulated by

@@

\section{Causal discovery methodology}
% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since the partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.} directly from a symmetric adjacency matrix.

By not taking into account the directivity for directed links, we systematically 
make a true positive and a false positive for such links (see a more detailed 
discussion about the chosen metric in section \ref{sec:metric}). Therefore, 
performance might be improved by predicting an asymmetric adjacency matrix.  
Hoping to orientate correctly the most obvious directed links, we attempt to 
retrieve some causal information, i.e. the directivity of the links, by computing 
and stacking with our symmetric adjacency matrix a matrix of pairwise activation 
heuristics. This method computes an anti-symmetric activation score $z_{ij}$ 
based the variation of fluorescence signal of a neuron $j$ due to a neuron $i$ 
over all timesteps:
\[
z_{ij} = \sum_{t=1}^n I(X_{t+1,j} - X_{t, i} \in [f_1, f_2]) -  I(X_{t+1,i} - X_{t, j} \in [f_1, f_2])
\]
where $f_1$ and $f_2$ are parameters of the method and $I$ is the
indicator function. % TODO say that we use only the 3 first filters for this

@@

\end{tabular}
\end{table}
\subsection{Discussion about the metric}
\label{sec:metric}
